---
title: "CSDA1040_Lab1"
author: "Aaron"
date: "9/16/2020"
output: pdf_document
---

### Load the libraries
```{r, include=FALSE}
library(data.table)
library(dplyr)
library(ggplot2)
library(knitr)
library(stringr)
library(DT)
library(gridExtra)
library(tidyverse)
library(recommenderlab)
library(tidyfst)
library(treemap)
```

### Load the data
```{r data, include=FALSE, echo=FALSE}
orders <- fread("C:/Users/axiao/Downloads/data/instacart-market-basket-analysis/orders.csv")
products <- fread('C:/Users/axiao/Downloads/data/instacart-market-basket-analysis/products.csv')
order_products <- fread('C:/Users/axiao/Downloads/data/instacart-market-basket-analysis/order_products__train.csv')
order_products_prior <- fread('C:/Users/axiao/Downloads/data/instacart-market-basket-analysis/order_products__prior.csv')
aisles <- fread('C:/Users/axiao/Downloads/data/instacart-market-basket-analysis/aisles.csv')
departments <- fread('C:/Users/axiao/Downloads/data/instacart-market-basket-analysis/departments.csv')
```

### Load the data
```{r data, include=FALSE, echo=FALSE}
orders_csv <- readr::read_csv("C:/Users/axiao/Downloads/data/instacart-market-basket-analysis/orders.csv")
# products <- fread('C:/Users/axiao/Downloads/data/instacart-market-basket-analysis/products.csv')
# order_products <- fread('C:/Users/axiao/Downloads/data/instacart-market-basket-analysis/order_products__train.csv')
# order_products_prior <- fread('C:/Users/axiao/Downloads/data/instacart-market-basket-analysis/order_products__prior.csv')
# aisles <- fread('C:/Users/axiao/Downloads/data/instacart-market-basket-analysis/aisles.csv')
# departments <- fread('C:/Users/axiao/Downloads/data/instacart-market-basket-analysis/departments.csv')
```

```{r}
head(orders_csv)
head(orders)
```

### Peek at the dataset 1
```{r}
# kable(head(orders,10))
# kable(head(products,10))
# kable(head(order_products,10))
# kable(head(order_products_prior,10))
# kable(head(aisles,10))
# kable(head(departments,10))
```

### Peek at the dataset 2
```{r}
# str(orders)
# str(products)
# str(order_products)
# str(order_products_prior)
# str(aisles)
# str(departments)
```

```{r}
# colSums(is.na(orders))
# colSums(is.na(products))
# colSums(is.na(order_products))
# colSums(is.na(order_products_prior))
# colSums(is.na(aisles))
# colSums(is.na(departments))
```

### Recode variables
```{r}
orders <- orders %>% mutate(order_hour_of_day = as.numeric(order_hour_of_day), 
                            eval_set = as.factor(eval_set))
products <- products %>% mutate(product_name = as.factor(product_name))
aisles <- aisles %>% mutate(aisle = as.factor(aisle))
departments <- departments %>% mutate(department = as.factor(department))
```

## The order behaivor

### Hour of Day
```{r}
orders %>% 
  ggplot(aes(x=order_hour_of_day)) + 
  geom_histogram(stat="count",fill=rainbow(24))
```

### Day of Week
```{r}
orders %>% 
  ggplot(aes(x=order_dow)) + 
  geom_histogram(stat="count",fill=rainbow(7))
```

### When do they order again
```{r}
orders %>% 
  ggplot(aes(x=days_since_prior_order)) + 
  geom_histogram(stat="count",fill="blue")
```

### How many prior orders are there?
```{r}
prior_order = count(filter(orders, eval_set=="prior"), order_number)
ggplot(data = prior_order, aes(order_number,n)) +
  geom_line(color="blue", size=1) +
  geom_point(color="blue", size=2)
```

### The distributions of how many items are in the orders
```{r}
train_1 <- order_products %>% 
          group_by(order_id) %>% 
          summarize(n_items_train = last(add_to_cart_order))
prior_1 <- order_products_prior %>% 
          group_by(order_id) %>% 
          summarize(n_items_prior = last(add_to_cart_order))

ggplot() + 
  geom_histogram(data=train_1, aes(x=n_items_train),col="red",alpha = 0.3) +
  geom_histogram(data=prior_1, aes(x=n_items_prior),col="blue",alpha = 0.3) +
  coord_cartesian(xlim=c(0,80))
```

### How often do people order the same items again
```{r}
re_order <- order_products %>% 
            group_by(reordered) %>% 
            summarize(count = n()) %>% 
  mutate(reordered = as.factor(reordered)) %>%
  mutate(proportion = count/sum(count))
head(re_order)
```

```{r}
ggplot(re_order, aes(x=reordered,y=count,fill=reordered))+
  geom_bar(stat="identity")
```

### Most Popular Products Sold
```{r}
tmp1 <- order_products %>%
  left_join(products) %>%
  group_by(product_name) %>%
  summarize(count=n()) %>%
  top_n(n=20, wt=count) %>%  
  mutate(percentage=count/sum(count))
p1 = ggplot (tmp1, aes(x=reorder(product_name,count), y=percentage)) +  
  geom_col() + 
  ggtitle('Products Top 20') + 
  ylab('Percentage of Orders') +
  theme (axis.text.x=element_text(angle=90, hjust=1, vjust=0.5),
         axis.title.x = element_blank()) 
p2 = ggplot (data = tmp1, aes( x= '', y=percentage )) + 
  ggtitle('Products Top 20') + 
  ylab('percentage.of.orders') + 
  geom_boxplot() + 
  xlab('Products')
grid.arrange(p1, p2, ncol = 2)
```

### Most Popular Department Sold
```{r}
tmp2 <- order_products %>%
  left_join(products) %>%
  left_join(departments) %>%
  group_by(department) %>%
  summarize(count=n()) %>%
  mutate(percentage=count/sum(count))
p1 = ggplot (tmp2, aes(x=reorder(department,count), y=percentage)) +  
  geom_col() + 
  ggtitle('Departments') + 
  ylab('Percentage of Orders') +
  theme (axis.text.x=element_text(angle=90, hjust=1, vjust=0.5),
         axis.title.x = element_blank()) 
p2 = ggplot (data = tmp2, aes( x= '', y=percentage )) + 
  ggtitle('Departments') + 
  ylab('percentage.of.orders') + 
  geom_boxplot() + 
  xlab('Departments')
grid.arrange(p1, p2, ncol = 2)
```

### Most Popular Aisles Sold
```{r}
tmp3 <- order_products %>%
  left_join(products) %>%
  left_join(aisles) %>%
  group_by(aisle) %>%
  summarize(count=n()) %>%
  top_n(n=20, wt=count) %>%  
  mutate(percentage=count/sum(count))
p5 = ggplot (tmp3, aes(x=reorder(aisle,count), y=percentage)) +  
  geom_col() + 
  ggtitle('Aisles Top 20') + 
  ylab('Percentage of Orders') +
  theme (axis.text.x=element_text(angle=90, hjust=1, vjust=0.5),
         axis.title.x = element_blank()) 
p6 = ggplot (data = tmp3, aes( x= '', y=percentage )) + 
  ggtitle('Aisles Top 20') + 
  ylab('percentage.of.orders') + 
  geom_boxplot() + 
  xlab('Products')
grid.arrange(p5, p6, ncol = 2)
```

### Top ten products ordered daily contributes between 7% to 8%.
```{r}
order_products_prior %>% 
  left_join(orders) %>% left_join(products) %>%
  group_by(order_dow, product_name) %>%
  summarize(n=n()) %>%
  mutate(percentage=n/sum(n)) %>%
  top_n(10, wt=n) %>%
  ggplot (aes(x=as.factor(order_dow), y=percentage, fill=product_name)) + 
    geom_col() + ylab('Proprtion of Orders') + ggtitle('Daily Top 10 Products Ordered') +
    theme(legend.position="bottom",legend.direction="horizontal")
```

### 
```{r}
order_products_prior %>% 
  left_join(orders) %>% left_join(products) %>%
  group_by(order_hour_of_day, product_name) %>%
  summarize(n=n()) %>%
  mutate(percentage=n/sum(n)) %>%
  top_n(10, wt=n) %>%
  ggplot(aes(x=as.factor(order_hour_of_day), y=percentage, fill=product_name)) + 
    geom_col() + ylab('Proprtion of Orders In An Hour') + 
    ggtitle('Hourly Top 10 Products Ordered') +
    theme(legend.position="bottom",legend.direction="horizontal")
```

## Visualizing the Product Portfolio

### use treemap package to visualize the structure of instacarts product portfolio, 
```{r}
tmp4 <- products %>% 
  group_by(department_id, aisle_id) %>% 
  summarize(n=n()) %>% 
  left_join(departments, by="department_id") %>% 
  left_join(aisles, by="aisle_id")
```

### 
```{r}
tmp5 <-order_products %>% 
  group_by(product_id) %>% 
  summarize(count=n()) %>% 
  left_join(products, by="product_id") %>% 
  ungroup() %>% 
  group_by(department_id, aisle_id) %>% 
  summarize(sumcount = sum(count)) %>% 
  left_join(tmp4, by = c("department_id", "aisle_id")) %>% 
  mutate(onesize = 1)
```

### Visualize in aisles organized within departments?
```{r}
treemap(tmp5,index=c("department","aisle"), vSize="onesize",vColor="department", 
        palette="Set3", title="", sortID="-sumcount", border.col="#FFFFFF") 
        #type="categorical", fontsize.legend = 0,bg.labels = "#FFFFFF")
```

### How many unique products are offered in each department/aisle?
```{r}
# The size of the boxes shows the number of products in each category.
treemap(tmp5,index=c("department","aisle"),vSize="n",title="",palette="Set3",border.col="#FFFFFF")
```

### How often are products from the department/aisle sold?
```{r}
# The size of the boxes shows the number of sales.
treemap(tmp5,index=c("department","aisle"),vSize="sumcount",title="",palette="Set3",border.col="#FFFFFF")
```


## Predictive Analysis

### Create a training label, which is 1 or 0, to indicate the actual basket content.
```{r}
data_train = orders %>% 
  filter(eval_set=='train') %>% 
  left_join(order_products) %>%
  left_join(products) %>%
  mutate(actual=1) %>%   #this is training label
  select(user_id, order_id, product_id, product_name, actual)
```

###
```{r}
# data_train <- data_train %>% mutate(actual = as.double(actual),
#                                     user_id = as.double(user_id))
glimpse(data_train)
head(data_train)
```

###
```{r}
summary(data_train)
```

###
```{r}
train01 <- data_train %>% 
  filter(user_id <= 25000)
glimpse(train01)
```

###
```{r}

```

###
```{r}
ratings_matrix <- train01 %>%
# Select only needed variables
  select(user_id, product_name) %>% 
# Add a column of 1s
  mutate(value = 1) %>%
# Spread into user-item format
  spread(product_name, value, fill = 0) %>% #, convert = FALSE, drop = TRUE
  select(-user_id) %>%
# Convert to matrix
  as.matrix() %>%
# Convert to recommenderlab class 'binaryRatingsMatrix'
  as("binaryRatingMatrix")
ratings_matrix
```

###
```{r}
train01 %>% 
  select(user_id, product_name) %>% 
  mutate(n=1)  %>%
  arrange(product_name) %>% 
  pivot_wider(names_from = "product_name", values_from = "n", values_fill=0)
```

###
```{r}
scheme <- ratings_matrix %>% 
  evaluationScheme(method = "cross",
                   k      = 5, 
                   train  = 0.7,  
                   given  = -1)
scheme
```

###
```{r}
algorithms <- list(
  "association rules" = list(name  = "AR", 
                        param = list(supp = 0.01, conf = 0.01)),
  "random items"      = list(name  = "RANDOM",  param = NULL),
  "popular items"     = list(name  = "POPULAR", param = NULL),
  "item-based CF"     = list(name  = "IBCF", param = list(k = 5)),
  "user-based CF"     = list(name  = "UBCF", 
                        param = list(method = "Cosine", nn = 500))
                   )
```

###
```{r}

```

###
```{r}

```

###
```{r}

```

###
```{r}

```

###
```{r}

```

###
```{r}

```

###
```{r}

```

###
```{r}

```

###
```{r}

```

###
```{r}

```
